---
title: "Alternative predictors"
description: "Example 6.3 of Mize, Doan, and Long (2019)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alternative predictors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.3 of Mize, Doan, and Long (2019).

This example uses two binary logit models with alternative predictors: two
different ways to measure sexuality, sexual behavior (`sexbehav`) in one model
and sexual identity (`sexident`) in the other.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-3-models}
vars63 <- c("samesexB", "sexident", "sexbehav", "college",
            "woman", "race", "age", "year")
dat63 <- gss[complete.cases(gss[vars63]), ]
nrow(dat63)
dat63 <- factorize(dat63, c("sexident", "sexbehav", "college", "woman", "race", "year"))

behavior63 <- glm(samesexB ~ sexbehav + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
identity63 <- glm(samesexB ~ sexident + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
fit63 <- suest(behavior63, identity63, model_names = c("Behavior", "Identity"))

effects63 <- avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"), newdata = dat63)
effects63
```

Because each predictor is in only one of the models, `marginaleffects`
reports zero and `NA` for the effects that don't exist in a model. The nonzero
rows are the effects for each measure.

```{r example-6-3-contrasts}
alternative_predictor_hypothesis <- function(x) {
  group <- as.character(x$group)
  term <- as.character(x$term)
  behavior <- x$estimate[term == "sexbehav" & group == "Behavior"]
  identity <- x$estimate[term == "sexident" & group == "Identity"]

  data.frame(
    term = c("Behavior bisexual - Identity bisexual",
             "Behavior gay - Identity gay",
             "(Behavior gay - bisexual) - (Identity gay - bisexual)"),
    estimate = c(behavior[1] - identity[1], behavior[2] - identity[2],
                 (behavior[2] - behavior[1]) - (identity[2] - identity[1])))
}

avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"),
  newdata = dat63, hypothesis = alternative_predictor_hypothesis)
```

**Example interpretations:** The contrasts between heterosexual and
gay/lesbian respondents and between bisexual and gay/lesbian respondents do
not differ significantly across the two models. However, the difference
between heterosexual and bisexual respondents is significantly larger when
sexuality is measured using identity than when it is measured using behavior.
