---
title: "Different outcomes"
description: "Example 6.4 of Mize, Doan, and Long (2019)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Different outcomes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.4 of Mize, Doan, and Long (2019). The
[same example in Stata](https://www.trentonmize.com/software/mecompare/examples/different-outcomes.html)
uses `mecompare`.

This example compares effects across different outcomes. Two counts, poor
mental-health days and poor physical-health days, are modeled with negative
binomial regression, and effects are on the predicted-rate scale. Effects and
cross-model comparisons are calculated for all predictors. The sample also
drops cases missing `reltrad`, which keeps the analysis sample of the paper
even though `reltrad` isn't in either model.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-4}
vars64 <- c("mntlhlth", "physhlth", "woman", "married", "age",
            "faminc", "race", "college", "parent", "reltrad")
dat64 <- gss[complete.cases(gss[vars64]), ]
stopifnot(nrow(dat64) == 5062)
nrow(dat64)

dat64 <- factorize(dat64, c("woman", "married", "parent", "college", "race", "year"))

mental64 <- MASS::glm.nb(mntlhlth ~ woman + married + parent + college + age +
                           faminc + race + year, data = dat64)
physical64 <- MASS::glm.nb(physhlth ~ woman + married + parent + college + age +
                             faminc + race + year, data = dat64)
fit64 <- suest(mental64, physical64, model_names = c("Mental", "Physical"))

variables64 <- list(woman = "reference", married = "reference", parent = "reference",
  college = "reference", age = sd(dat64$age), faminc = sd(dat64$faminc),
  race = "reference", year = "reference")
effects64 <- avg_comparisons(fit64, variables = variables64, newdata = dat64)
effects64

age64 <- avg_comparisons(fit64, variables = list(age = sd(dat64$age)),
                         newdata = dat64)
hypotheses(age64, hypothesis = difference ~ revpairwise)
```

`age = sd(dat64$age)` asks for a one-standard-deviation increase from each
person's observed age.

**Example interpretations:** Women report about 0.99 more poor
mental-health days and 0.77 more poor physical-health days per month than men.
Although the effect of gender is about 0.22 larger for mental health, the
cross-outcome difference is not statistically significant.

Being married significantly reduces poor mental-health days by about 1.01,
whereas its effect on poor physical-health days is not statistically
significant. The direct comparison shows that the effect of marriage is
significantly larger for mental health than for physical health
(cross-model difference = -0.851, \(p < .01\)).

Similarly, the effect of age differs significantly across the outcomes:
aging is associated with fewer poor mental-health days but more poor
physical-health days.
