---
title: "Example 6.4: Different count outcomes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example 6.4: Different count outcomes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


These data and specifications reproduce the corresponding example in Mize,
Doan, and Long (2019) and on the
[`mecompare` documentation page](https://www.trentonmize.com/software/mecompare).
The code chunks are not evaluated while building the package website because
they download the public replication data. Copy and run the code interactively
to reproduce the results.


## Load and prepare the data

```{r}
library(haven)
library(MASS)
library(marginaleffects)
library(suest)

gss <- zap_labels(read_dta(
  "https://tdmize.github.io/data/data/gss_cme.dta"
))

vars <- c(
  "mntlhlth", "physhlth", "woman", "married", "age",
  "faminc", "race", "college", "parent", "year"
)
dat <- gss[complete.cases(gss[vars]), ]
dat[c("woman", "married", "parent", "college", "race", "year")] <-
  lapply(dat[c("woman", "married", "parent", "college",
               "race", "year")], factor)
```

## Fit negative-binomial models for two outcomes

```{r}
mental <- glm.nb(
  mntlhlth ~ woman + married + parent + college + age +
    faminc + race + year,
  data = dat
)

physical <- glm.nb(
  physhlth ~ woman + married + parent + college + age +
    faminc + race + year,
  data = dat
)

combined <- suest(
  mental,
  physical,
  model_names = c("Mental", "Physical")
)
```

`suest()` includes `log(theta)` from each negative-binomial model in the joint
parameter vector, which is required to reproduce the robust uncertainty
estimates.

## Compare effects across outcomes

```{r}
forward_change <- function(amount) {
  function(x) data.frame(lo = x, hi = x + amount)
}

effects <- avg_comparisons(
  combined,
  variables = list(
    woman = "reference",
    married = "reference",
    parent = "reference",
    college = "reference",
    age = forward_change(sd(dat$age)),
    faminc = forward_change(sd(dat$faminc)),
    race = "reference",
    year = "reference"
  ),
  newdata = dat
)

effects
```

For example, the one-standard-deviation age effects are approximately -0.462
poor mental-health days and 0.491 poor physical-health days. Their difference
is approximately -0.953 with a robust standard error of 0.132.
