---
title: "Example 6.5: Ordered versus multinomial logit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example 6.5: Ordered versus multinomial logit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


These data and specifications reproduce the corresponding example in Mize,
Doan, and Long (2019) and on the
[`mecompare` documentation page](https://www.trentonmize.com/software/mecompare).
The code chunks are not evaluated while building the package website because
they download the public replication data. Copy and run the code interactively
to reproduce the results.


## Load and prepare the data

```{r}
library(haven)
library(MASS)
library(nnet)
library(marginaleffects)
library(suest)

gss <- zap_labels(read_dta(
  "https://tdmize.github.io/data/data/gss_cme.dta"
))

vars <- c(
  "partyid5", "woman", "edyrs", "age", "parent", "married",
  "faminc", "employed", "region4", "year", "race"
)

dat <- subset(gss, year >= 2010)
dat <- dat[complete.cases(dat[vars]), ]
dat[c("woman", "parent", "married", "race",
      "employed", "region4", "year")] <-
  lapply(dat[c("woman", "parent", "married", "race",
               "employed", "region4", "year")], factor)

party_levels <- sort(unique(dat$partyid5))
dat$party_ord <- ordered(dat$partyid5, levels = party_levels)
dat$party_nom <- factor(dat$partyid5, levels = party_levels)
```

## Fit and combine the models

```{r}
ordered_model <- polr(
  party_ord ~ age + I(age^2) + woman + edyrs + parent +
    married + race + faminc + employed + region4 + year,
  data = dat,
  method = "logistic",
  Hess = TRUE
)

nominal_model <- multinom(
  party_nom ~ age + I(age^2) + woman + edyrs + parent +
    married + race + faminc + employed + region4 + year,
  data = dat,
  Hess = TRUE,
  trace = FALSE
)

combined <- suest(
  ordered_model,
  nominal_model,
  model_names = c("Ordered", "Multinomial")
)
```

## Compare category probabilities

For a standard empirical-distribution average, `marginaleffects` can calculate
category-specific changes directly:

```{r}
avg_comparisons(
  combined,
  variables = list(age = function(x) {
    data.frame(lo = x, hi = x + 10)
  }),
  newdata = dat
)
```

The published example evaluates a change from age 20 to age 30 with all other
model-matrix columns held at their means. The package's full acceptance suite
implements that exact delta-method calculation and reproduces the Stata
results:

| Category | Ordered effect | Ordered SE | Nominal effect | Nominal SE | Difference | Difference SE |
|---|---:|---:|---:|---:|---:|---:|
| Strong Democrat | 0.020 | 0.004 | 0.032 | 0.003 | -0.012 | 0.003 |
| Democrat | 0.023 | 0.005 | -0.010 | 0.011 | 0.034 | 0.009 |
| Independent | -0.003 | 0.000 | -0.000 | 0.010 | -0.003 | 0.010 |
| Republican | -0.024 | 0.005 | -0.031 | 0.011 | 0.007 | 0.009 |
| Strong Republican | -0.016 | 0.004 | 0.010 | 0.003 | -0.026 | 0.004 |
