---
title: "Example 6.2: Nested logit models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example 6.2: Nested logit models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


These data and specifications reproduce the corresponding example in Mize,
Doan, and Long (2019) and on the
[`mecompare` documentation page](https://www.trentonmize.com/software/mecompare).
The code chunks are not evaluated while building the package website because
they download the public replication data. Copy and run the code interactively
to reproduce the results.


## Load and prepare the data

```{r}
library(haven)
library(marginaleffects)
library(suest)

gss <- zap_labels(read_dta(
  "https://tdmize.github.io/data/data/gss_cme.dta"
))

vars <- c(
  "vhappy", "college", "age", "married", "parent", "woman",
  "conserv", "reltrad", "year", "employed"
)

dat <- subset(gss, year >= 2000 & employed == 1)
dat <- dat[complete.cases(dat[vars]), ]
dat[c("college", "married", "parent", "woman", "conserv",
      "reltrad", "year")] <-
  lapply(dat[c("college", "married", "parent", "woman", "conserv",
               "reltrad", "year")], factor)
```

## Fit and combine the models

```{r}
model1 <- glm(
  vhappy ~ college,
  family = binomial("logit"),
  data = dat
)

model2 <- glm(
  vhappy ~ college + married + parent + woman + conserv +
    reltrad + year + age + I(age^2),
  family = binomial("logit"),
  data = dat
)

combined <- suest(
  model1,
  model2,
  model_names = c("Model 1", "Model 2")
)
```

## Compare the average effect of college

```{r}
effects <- avg_comparisons(
  combined,
  variables = "college",
  newdata = dat
)

effects
hypotheses(effects, hypothesis = difference ~ revpairwise)
```

The expected average effects are about 0.0718 and 0.0599. The cross-model
difference is about 0.0119 with a robust standard error of 0.0040.
