---
title: "Example 6.3: Alternative predictor operationalizations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example 6.3: Alternative predictor operationalizations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


These data and specifications reproduce the corresponding example in Mize,
Doan, and Long (2019) and on the
[`mecompare` documentation page](https://www.trentonmize.com/software/mecompare).
The code chunks are not evaluated while building the package website because
they download the public replication data. Copy and run the code interactively
to reproduce the results.


## Load and prepare the data

```{r}
library(haven)
library(marginaleffects)
library(suest)

gss <- zap_labels(read_dta(
  "https://tdmize.github.io/data/data/gss_cme.dta"
))

vars <- c(
  "samesexB", "sexident", "sexbehav", "college",
  "woman", "race", "age", "year"
)
dat <- gss[complete.cases(gss[vars]), ]
dat[c("sexident", "sexbehav", "college", "woman", "race", "year")] <-
  lapply(dat[c("sexident", "sexbehav", "college", "woman",
               "race", "year")], factor)
```

## Fit models using behavior and identity

```{r}
behavior <- glm(
  samesexB ~ sexbehav + woman + college + age + race + year,
  family = binomial("logit"),
  data = dat
)

identity <- glm(
  samesexB ~ sexident + woman + college + age + race + year,
  family = binomial("logit"),
  data = dat
)

combined <- suest(
  behavior,
  identity,
  model_names = c("Behavior", "Identity")
)
```

## Estimate the category effects

```{r}
effects <- avg_comparisons(
  combined,
  variables = list(
    sexbehav = "reference",
    sexident = "reference"
  ),
  newdata = dat
)

effects
```

Because each predictor appears in only one component model,
`marginaleffects` reports zero and `NA` for the structurally absent effects.
The nonzero rows give the effects for each operationalization.

## Test custom cross-model contrasts

```{r}
alternative_predictor_hypothesis <- function(x) {
  group <- as.character(x$group)
  term <- as.character(x$term)

  behavior <- x$estimate[term == "sexbehav" & group == "Behavior"]
  identity <- x$estimate[term == "sexident" & group == "Identity"]

  data.frame(
    term = c(
      "Behavior bisexual - Identity bisexual",
      "Behavior gay - Identity gay",
      "(Behavior gay - bisexual) - (Identity gay - bisexual)"
    ),
    estimate = c(
      behavior[1] - identity[1],
      behavior[2] - identity[2],
      (behavior[2] - behavior[1]) -
        (identity[2] - identity[1])
    )
  )
}

avg_comparisons(
  combined,
  variables = list(
    sexbehav = "reference",
    sexident = "reference"
  ),
  newdata = dat,
  hypothesis = alternative_predictor_hypothesis
)
```

The expected cross-model contrasts are approximately 0.177, 0.066, and
-0.111.
