---
title: "Separate samples or groups"
description: "Example 6.6 of Mize, Doan, and Long (2019)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Separate samples or groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.6 of Mize, Doan, and Long (2019).

This example compares effects from the same binary logit model fitted to two
different samples, one from 1986 and one from 2016. The same approach works for
comparing groups, such as men and women. The samples are selected with
`subset`, and `suest_newdata()` makes sure each marginal effect is averaged
over its own model's sample.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-6}
vars66 <- c("helpsickB", "polviews", "conserv", "faminc", "employed",
            "woman", "age", "college", "married", "parent", "race", "year")
dat66 <- gss[complete.cases(gss[vars66]), ]
dat66 <- factorize(dat66, c("conserv", "employed", "woman", "college",
                            "married", "parent", "race"))

model1986 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 1986)
model2016 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 2016)
c(`1986` = nobs(model1986), `2016` = nobs(model2016))
fit66 <- suest(model1986, model2016, model_names = c("1986", "2016"))
fit66

effects66 <- avg_comparisons(fit66, variables = "conserv",
                             newdata = suest_newdata(fit66))
effects66
hypotheses(effects66, hypothesis = difference ~ revpairwise)
```

**Example interpretation:** Conservatives had significantly lower predicted
probabilities of saying that government should be responsible for providing
health care in both 1986 (marginal effect = -0.092) and 2016 (marginal effect =
-0.258). The gap between conservatives and nonconservatives increased over
time: the conservative effect is significantly more negative in 2016 than in
1986 (cross-model difference = 0.166, \(p < .001\)).
