---
title: "Curvilinear effects and mediation/attenuation"
description: "Example 6.1 of Mize, Doan, and Long (2019)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Curvilinear effects and mediation/attenuation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.1 of Mize, Doan, and Long (2019). The
[same example in Stata](https://www.trentonmize.com/software/mecompare/examples/mediation.html)
uses `mecompare`.

This example uses a linear regression with a nonlinear effect of income,
included as income and income squared. We test whether adding a mediator, job
satisfaction, reduces the average effect of income, which would suggest
mediation. It also shows how to ask for a custom amount of change in a
continuous predictor: here, a one-standard-deviation increase in income.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

ah <- zap_labels(read_dta("https://tdmize.github.io/data/data/ah4_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-1}
vars61 <- c("depsympB", "income", "inc10", "age", "woman", "race", "college", "jobsat")
dat61 <- ah[complete.cases(ah[vars61]), ]
stopifnot(nrow(dat61) == 4307)
nrow(dat61)
dat61 <- factorize(dat61, c("woman", "race", "jobsat"))

base61 <- lm(depsympB ~ income + I(income^2) + age + woman + race, data = dat61)
mediator61 <- lm(depsympB ~ income + I(income^2) + age + woman + race + jobsat,
                 data = dat61)
fit61 <- suest(base61, mediator61, model_names = c("Base", "Mediator"))

effects61 <- avg_comparisons(fit61,
  variables = list(income = sd(dat61$income)), newdata = dat61)
effects61
hypotheses(effects61, hypothesis = difference ~ revpairwise)
```

A single number in `variables` asks for a change of that amount from each
observed value. So `list(income = sd(dat61$income))` compares each person's
observed income with that income plus one standard deviation. This is
different from `"sd"`, which asks for a one-standard-deviation change
centered on the mean.

**Example interpretation for the effect of income:** In model 1, the
marginal effect is -0.816, indicating that a standard-deviation increase in
income is associated with about 0.8 fewer depressive symptoms. The effect is
reduced to -0.648 after accounting for job satisfaction. The direct
cross-model test shows that accounting for job satisfaction decreases the
average effect of income by 0.168, a statistically significant reduction
(\(p < .001\)) that is consistent with mediation.
