---
title: "Different model types"
description: "Example 6.5 of Mize, Doan, and Long (2019): ordinal versus nominal."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Different model types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.5 of Mize, Doan, and Long (2019).

This example compares effects across two model types, an ordinal model and a
nominal model, with the same outcome and predictors. It also shows a custom
effect for a continuous predictor: a change in age from 20 to 30, with the
other covariates held at their means.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-5-models}
vars65 <- c("partyid5", "woman", "edyrs", "age", "parent", "married",
            "faminc", "employed", "region4", "year", "race")
dat65 <- subset(gss, year >= 2010)
dat65 <- dat65[complete.cases(dat65[vars65]), ]
nrow(dat65)
dat65 <- factorize(dat65, c("woman", "parent", "married", "race",
                            "employed", "region4", "year"))

party_levels <- sort(unique(dat65$partyid5))
party_labels <- c("Strong Democrat", "Democrat", "Independent",
                  "Republican", "Strong Republican")
dat65$party_ord <- ordered(dat65$partyid5, levels = party_levels, labels = party_labels)
dat65$party_nom <- factor(dat65$partyid5, levels = party_levels, labels = party_labels)

ordered65 <- MASS::polr(party_ord ~ age + I(age^2) + woman + edyrs + parent +
                          married + race + faminc + employed + region4 + year,
                        data = dat65, method = "logistic", Hess = TRUE)
nominal65 <- nnet::multinom(party_nom ~ age + I(age^2) + woman + edyrs + parent +
                             married + race + faminc + employed + region4 + year,
                           data = dat65, Hess = TRUE, trace = FALSE)
fit65 <- suest(ordered65, nominal65, model_names = c("Ordered", "Multinomial"))
```

The effect of age compares age 20 with age 30, with every other column of the
model matrix held at its mean (including each category of the factor
predictors). This is not the same as averaging a ten-year increase over
everyone's observed age and covariates, and the difference matters a lot here
because age is included as a quadratic. `newdata = "mean"` in
`marginaleffects` is different too: it uses means for numeric variables but
the most common category for factors. A helper included with `suest` does the
calculation. [View the helper code](https://github.com/tdmize/suest/blob/main/inst/example-code/example-6-5-atmeans.R).

```{r example-6-5-effects}
source(system.file("example-code", "example-6-5-atmeans.R", package = "suest"))
results65 <- example65_atmeans(fit65, age_lo = 20, age_hi = 30)
results65$effects
results65$differences
```

**Example interpretation:** For someone who is 20 years old, the effect of a
ten-year increase in age differs significantly across the ordinal and nominal
models for three of the five outcome categories. For example, aging increases
the probability of identifying as a strong Democrat in both models, but the
increase is significantly larger in the nominal model. More strikingly, the
effects of age on identifying as a strong Republican run in opposite
directions across the two models.
