---
title: "Notes on specific models"
description: "Settings and limits for a few of the supported models."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Notes on specific models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Most supported models need nothing special. The notes below cover the few
that do. Fits that `suest` can't use, such as penalized fits or models that
did not converge, stop with an error that explains why.

## Linear models

Linear models add an `lnvar` parameter (the log of the error variance) to the
combined coefficients and covariance matrix. It has no effect on predictions
or marginal effects.

## Linear panel models

For `plm::plm()` models, comparisons of slopes and changes are supported.
Confidence intervals and tests for average predicted levels at the
estimation-sample means are not: their standard errors can be missing or near
zero.

## Mixed models fitted with glmmTMB

- Predictions and marginal effects average over the random intercept, and
  over the random slope when there is one.
- Center continuous predictors and keep them in well-scaled units. glmmTMB's
  covariance matrix can be inaccurate with extreme predictor units even when
  the model reports convergence, and `suest` uses that covariance as supplied.
- For slopes of continuous predictors in logit random-slope models, use
  `numderiv = list("fdcenter", eps = 1e-4)` in `avg_slopes()` for more stable
  standard errors, and check that nearby step sizes (such as `5e-5` and
  `2e-4`) give similar answers.

## Models fitted with different packages

Models fitted by different packages can be combined, as long as they are
supported models.

```{r libraries}
library(suest)
library(marginaleffects)
```

`glm()` and `glm2::glm2()`:

```{r glm2-engine}
library(glm2)

dat <- mtcars
dat$am <- factor(dat$am)

standard <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
alternative <- glm2(am ~ wt + hp, family = binomial("logit"), data = dat)
engine_fit <- suest(standard, alternative, model_names = c("glm", "glm2"))
avg_comparisons(engine_fit, variables = "wt", newdata = dat)
```

`ordinal::clm()` and `MASS::polr()`. `clm()` models need flexible thresholds
and no scale or nominal-effects formula.

```{r clm-engine}
library(ordinal)

housing <- MASS::housing[rep(seq_len(nrow(MASS::housing)), MASS::housing$Freq),
                         c("Sat", "Infl", "Type", "Cont")]
rownames(housing) <- NULL
housing$Sat <- ordered(housing$Sat, levels = c("Low", "Medium", "High"))

clm_model <- clm(Sat ~ Cont + Infl + Type, data = housing, link = "logit")
polr_model <- MASS::polr(Sat ~ Cont + Infl + Type, data = housing,
                         method = "logistic", Hess = TRUE)
ordinal_fit <- suest(clm_model, polr_model, model_names = c("clm", "polr"))
avg_comparisons(ordinal_fit, variables = "Cont", newdata = housing)
```
