---
title: "Nested logit models and mediation/attenuation"
description: "Example 6.2 of Mize, Doan, and Long (2019)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nested logit models and mediation/attenuation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

Example 6.2 of Mize, Doan, and Long (2019). The
[same example in Stata](https://www.trentonmize.com/software/mecompare/examples/nested-models.html)
uses `mecompare`.

This example uses binary logit models to see how the effect of college
changes when more variables are added to the model.

```{r data}
library(suest)
library(marginaleffects)
library(haven)

gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

```{r example-6-2}
vars62 <- c("vhappy", "college", "wages", "occprest", "age", "married",
            "parent", "woman", "conserv", "reltrad", "year", "employed")
dat62 <- subset(gss, year >= 2000 & employed == 1)
dat62 <- dat62[complete.cases(dat62[vars62]), ]
stopifnot(nrow(dat62) == 9216)
nrow(dat62)
dat62 <- factorize(dat62, c("college", "married", "parent", "woman", "conserv",
                            "reltrad", "year"))

model62a <- glm(vhappy ~ college, family = binomial("logit"), data = dat62)
model62b <- glm(vhappy ~ college + married + parent + woman + conserv + reltrad +
                  year + age + I(age^2), family = binomial("logit"), data = dat62)
fit62 <- suest(model62a, model62b, model_names = c("Model 1", "Model 2"))

effects62 <- avg_comparisons(fit62, variables = "college", newdata = dat62)
effects62
hypotheses(effects62, hypothesis = difference ~ revpairwise)
```

**Example interpretation for the effect of college:** On average, the
probability of being very happy is 0.072 higher for people with a college
degree than for those without one (\(p < .001\)). Model 2 adds demographic
controls, reducing the average marginal effect of college to 0.060, which
remains statistically significant. A direct test shows that adding the
controls decreases the effect of college by 0.012 (\(p < .01\)).
