---
title: "Getting started with suest"
description: "Compare predictions and marginal effects across regression models in R."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with suest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

`suest()` combines two or more separately fitted models so you can test
whether predictions, marginal effects, or other quantities differ across the
models. It supplies the joint robust covariance matrix those tests need, and
leaves calculating and presenting the quantities themselves to
[`marginaleffects`](https://marginaleffects.com/).

The package implements the framework in Mize, Doan, and Long (2019). The six
examples below replicate Examples 6.1–6.6 for the
[`mecompare` Stata command](https://www.trentonmize.com/software/mecompare).

One note on continuous predictors: a single number in `variables` asks for a
change of that amount from each observed value. So
`list(income = sd(dat61$income))` compares each person's observed income with
that income plus one standard deviation. This is different from `"sd"`, which
asks for a one-standard-deviation change centered on the mean.

## Functions at a glance

- `suest()` combines two or more fitted models and calculates their joint
  covariance matrix.
- `suest_newdata()` stacks the models' estimation samples so each model's
  effects can be averaged over its own sample.
- The usual `marginaleffects` functions, including `predictions()`,
  `avg_comparisons()`, `avg_slopes()`, `hypotheses()`, and the plotting
  functions, work with the combined object.

## Installation and setup

```{r installation, eval=FALSE}
remotes::install_github("tdmize/suest")
```

```{r libraries}
library(suest)
library(marginaleffects)
```

## Basic workflow

```{r basic-workflow}
dat <- mtcars
dat$am <- factor(dat$am)

base <- glm(am ~ wt, family = binomial("logit"), data = dat)
adjusted <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
combined <- suest(base, adjusted, model_names = c("Base", "Adjusted"))
combined

effects <- avg_comparisons(combined, variables = "wt", newdata = dat)
effects
hypotheses(effects, hypothesis = difference ~ revpairwise)
```

The last command tests the difference between the two effects directly. That
is different from checking whether one effect is statistically significant and
the other is not.

## Supported models

| Model | Supported functions |
|---|---|
| Linear regression | `stats::lm()` |
| Binary logit, probit, and complementary-log-log | `stats::glm()`, `glm2::glm2()` |
| Poisson regression | `stats::glm()`, `glm2::glm2()` |
| Other GLMs with supported links | `stats::glm()`, `glm2::glm2()` |
| Fractional response | quasi-binomial logit, probit, complementary-log-log, or user-supplied log-log |
| Negative binomial regression | `MASS::glm.nb()` |
| Parametric survival and censored regression | `survival::survreg()` with a common scale, including Gaussian interval regression |
| Beta regression | `betareg::betareg()` with logit, probit, complementary-log-log, or log-log mean links |
| Zero-inflated Poisson and negative binomial | `pscl::zeroinfl()` |
| Truncated Gaussian regression | `truncreg::truncreg()` |
| Direct Tobit regression | `censReg::censReg()`; predictions are the latent mean |
| Instrumental-variable 2SLS | unweighted `fixest::feols()` without absorbed fixed effects |
| Heteroskedastic binary probit and logit | `Rchoice::hetprob()` |
| Instrumental-variable probit ML | `Rchoice::ivpml()`; response predictions are average structural probabilities |
| Bivariate probit | `mvProbit::mvProbit()` with common regressors across equations and `finalHessian = TRUE` |
| Linear panel fixed, between, and random effects | unweighted `plm::plm()` with individual effects; unbalanced random-effects panels can differ slightly from Stata |
| Random-intercept Gaussian panel ML | unweighted `nlme::lme(method = "ML")`; variance components included |
| Gaussian random-intercept GLMM | unweighted `glmmTMB`; binomial-logit, Poisson-log, or NB2-log with constant dispersion |
| Gaussian random-intercept and numeric-slope GLMM | unweighted `glmmTMB`; binomial-logit, Poisson-log, or NB2-log, `(1 + x | id)`, unstructured covariance; checked against Stata (see the README) |
| Population-averaged GEE | unweighted `geepack::geeglm()`; Gaussian, binary, or Poisson with independence/exchangeable correlation |
| Ordered logit and probit | `MASS::polr()`, restricted `ordinal::clm()` |
| Multinomial logit | `nnet::multinom()` |

Models can use the same, partially overlapping, or completely different
samples. Any supported models with a single outcome can be combined, and so
can models for categorical outcomes; the two kinds can also be mixed. Results
on different scales are labeled separately rather than treated as directly
comparable.

Panel and mixed (GLMM) models can currently be combined only with models of
the same type. For the random-slope mixed models, predictions average over
both random effects, and the joint covariance includes the random-effect
standard deviations and correlation (and, for NB2, the dispersion). Fit these
models with well-scaled predictors: glmmTMB's covariance matrix can be
inaccurate with extreme predictor units even when the model reports
convergence, and `suest` uses that covariance as supplied. For logit
random-slope models, slopes for continuous predictors also include changes in
the random-effect variance. For more stable numerical standard errors, use
`numderiv = list("fdcenter", eps = 1e-4)` in `marginaleffects::avg_slopes()`
and check that nearby step sizes give similar answers.

Offsets are supported. Bias-reduced, adjusted-score, Firth, and other
penalized GLM fits are rejected, because they are estimated differently from
ordinary maximum-likelihood GLMs.

As in Stata, linear models add an `lnvar` parameter (the log of the error
variance). Unweighted fits use `log(RSS / df.residual)`; weighted fits follow
`suest2`'s normalization. It has no effect on predictions or comparisons.

## Cluster-robust covariance

Supply a cluster variable to get cluster-robust standard errors for the
combined models:

```{r cluster-example, eval=FALSE}
fit_clustered <- suest(base, adjusted, cluster = "person_id")
avg_comparisons(fit_clustered, variables = "x", newdata = dat)
```

`cluster` can also be a list with one vector, matrix, or data frame matched
to each model's estimation sample. An observation that appears in more than
one model must have the same cluster ID in each. Observations in completely
different samples can belong to the same cluster, in which case the covariance
between models can be nonzero.

Panel models cluster on the panel identifier by default. A cluster you supply
must contain whole panels.

## Probability weights

`suest` supports probability weights for linear, binary logit and probit,
Poisson, negative binomial, ordered logit and probit, and multinomial logit
models. R's `weights=` argument doesn't say what kind of weights they are, so
you tell `suest()` explicitly.

```{r pweight-example}
set.seed(376)
weighted <- data.frame(x = rnorm(500), z = rnorm(500))
weighted$mediator <- 0.5 * weighted$x - 0.3 * weighted$z + rnorm(500)
weighted$pw <- exp(0.2 * weighted$z + 0.1 * rnorm(500))
weighted$y <- 1 + 0.7 * weighted$x - 0.4 * weighted$z +
  0.6 * weighted$mediator + rnorm(500)

base_w <- lm(y ~ x + z, weights = pw, data = weighted)
adjusted_w <- lm(y ~ x + z + mediator, weights = pw, data = weighted)
fit_w <- suest(base_w, adjusted_w, model_names = c("Base", "Adjusted"),
               weight_type = "pweight")
fit_w

effects_w <- avg_comparisons(fit_w, variables = "x", newdata = weighted, wts = "pw")
effects_w
hypotheses(effects_w, hypothesis = difference ~ revpairwise)
```

Weights must be positive and finite. An observation that appears in more than
one model must have the same weight in each; weights can differ for
observations in only one model, including completely different samples. When
the models have different estimation samples, `suest_newdata()` includes the
weights as `.suest_weight`:

```{r pweight-model-specific, eval=FALSE}
nd_w <- suest_newdata(fit_w)
avg_comparisons(fit_w, variables = "x", newdata = nd_w, wts = ".suest_weight")
```

For binary logit and probit, `suest` uses the observed information matrix,
which matches Stata's `suest` covariance.

## Replicating Mize, Doan, and Long (2019)

The examples use the public replication data for the
[`mecompare` Stata command](https://www.trentonmize.com/software/mecompare).
The data are downloaded once and used throughout.

```{r replication-data}
library(haven)

ah <- zap_labels(read_dta("https://tdmize.github.io/data/data/ah4_cme.dta"))
gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

### Ex 6.1 - Marginal effects to summarize curvilinear relationships and test mediation

This example uses a linear regression with a nonlinear effect of income,
included as income and income squared. We test whether adding a mediator, job
satisfaction, reduces the average effect of income, which would suggest
mediation. It also shows how to ask for a custom amount of change in a
continuous predictor: here, a one-standard-deviation increase in income.

```{r example-6-1}
vars61 <- c("depsympB", "income", "inc10", "age", "woman", "race", "college", "jobsat")
dat61 <- ah[complete.cases(ah[vars61]), ]
stopifnot(nrow(dat61) == 4307)
nrow(dat61)
dat61 <- factorize(dat61, c("woman", "race", "jobsat"))

base61 <- lm(depsympB ~ income + I(income^2) + age + woman + race, data = dat61)
mediator61 <- lm(depsympB ~ income + I(income^2) + age + woman + race + jobsat,
                 data = dat61)
fit61 <- suest(base61, mediator61, model_names = c("Base", "Mediator"))

effects61 <- avg_comparisons(fit61,
  variables = list(income = sd(dat61$income)), newdata = dat61)
effects61
hypotheses(effects61, hypothesis = difference ~ revpairwise)
```

**Example interpretation for the effect of income:** In model 1, the
marginal effect is -0.816, indicating that a standard-deviation increase in
income is associated with about 0.8 fewer depressive symptoms. The effect is
reduced to -0.648 after accounting for job satisfaction. The direct
cross-model test shows that accounting for job satisfaction decreases the
average effect of income by 0.168, a statistically significant reduction
(\(p < .001\)) that is consistent with mediation.

### Ex 6.2 - Comparing marginal effects across nested logit models

This example uses binary logit models to see how the effect of college
changes when more variables are added to the model.

```{r example-6-2}
vars62 <- c("vhappy", "college", "wages", "occprest", "age", "married",
            "parent", "woman", "conserv", "reltrad", "year", "employed")
dat62 <- subset(gss, year >= 2000 & employed == 1)
dat62 <- dat62[complete.cases(dat62[vars62]), ]
stopifnot(nrow(dat62) == 9216)
nrow(dat62)
dat62 <- factorize(dat62, c("college", "married", "parent", "woman", "conserv",
                            "reltrad", "year"))

model62a <- glm(vhappy ~ college, family = binomial("logit"), data = dat62)
model62b <- glm(vhappy ~ college + married + parent + woman + conserv + reltrad +
                  year + age + I(age^2), family = binomial("logit"), data = dat62)
fit62 <- suest(model62a, model62b, model_names = c("Model 1", "Model 2"))

effects62 <- avg_comparisons(fit62, variables = "college", newdata = dat62)
effects62
hypotheses(effects62, hypothesis = difference ~ revpairwise)
```

**Example interpretation for the effect of college:** On average, the
probability of being very happy is 0.072 higher for people with a college
degree than for those without one (\(p < .001\)). Model 2 adds demographic
controls, reducing the average marginal effect of college to 0.060, which
remains statistically significant. A direct test shows that adding the
controls decreases the effect of college by 0.012 (\(p < .01\)).

### Ex 6.3 - Comparing marginal effects using alternative predictors

This example uses two binary logit models with alternative predictors: two
different ways to measure sexuality, sexual behavior (`sexbehav`) in one model
and sexual identity (`sexident`) in the other.

```{r example-6-3-models}
vars63 <- c("samesexB", "sexident", "sexbehav", "college",
            "woman", "race", "age", "year")
dat63 <- gss[complete.cases(gss[vars63]), ]
stopifnot(nrow(dat63) == 4921)
nrow(dat63)
dat63 <- factorize(dat63, c("sexident", "sexbehav", "college", "woman", "race", "year"))

behavior63 <- glm(samesexB ~ sexbehav + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
identity63 <- glm(samesexB ~ sexident + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
fit63 <- suest(behavior63, identity63, model_names = c("Behavior", "Identity"))

effects63 <- avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"), newdata = dat63)
effects63
```

Because each predictor is in only one of the models, `marginaleffects`
reports zero and `NA` for the effects that don't exist in a model. The nonzero
rows are the effects for each measure.

```{r example-6-3-contrasts}
alternative_predictor_hypothesis <- function(x) {
  group <- as.character(x$group)
  term <- as.character(x$term)
  behavior <- x$estimate[term == "sexbehav" & group == "Behavior"]
  identity <- x$estimate[term == "sexident" & group == "Identity"]

  data.frame(
    term = c("Behavior bisexual - Identity bisexual",
             "Behavior gay - Identity gay",
             "(Behavior gay - bisexual) - (Identity gay - bisexual)"),
    estimate = c(behavior[1] - identity[1], behavior[2] - identity[2],
                 (behavior[2] - behavior[1]) - (identity[2] - identity[1])))
}

avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"),
  newdata = dat63, hypothesis = alternative_predictor_hypothesis)
```

**Example interpretations:** The contrasts between heterosexual and
gay/lesbian respondents and between bisexual and gay/lesbian respondents do
not differ significantly across the two models. However, the difference
between heterosexual and bisexual respondents is significantly larger when
sexuality is measured using identity than when it is measured using behavior.

### Ex 6.4 - Comparing marginal effects across different outcomes

This example compares effects across different outcomes. Two counts, poor
mental-health days and poor physical-health days, are modeled with negative
binomial regression, and effects are on the predicted-rate scale. Effects and
cross-model comparisons are calculated for all predictors. To match the
`mecompare` Stata example exactly, the sample also drops observations missing
`reltrad`, even though `reltrad` isn't in either model.

```{r example-6-4}
vars64 <- c("mntlhlth", "physhlth", "woman", "married", "age",
            "faminc", "race", "college", "parent", "reltrad")
dat64 <- gss[complete.cases(gss[vars64]), ]
stopifnot(nrow(dat64) == 5062)
nrow(dat64)

dat64 <- factorize(dat64, c("woman", "married", "parent", "college", "race", "year"))

mental64 <- MASS::glm.nb(mntlhlth ~ woman + married + parent + college + age +
                           faminc + race + year, data = dat64)
physical64 <- MASS::glm.nb(physhlth ~ woman + married + parent + college + age +
                             faminc + race + year, data = dat64)
fit64 <- suest(mental64, physical64, model_names = c("Mental", "Physical"))

variables64 <- list(woman = "reference", married = "reference", parent = "reference",
  college = "reference", age = sd(dat64$age), faminc = sd(dat64$faminc),
  race = "reference", year = "reference")
effects64 <- avg_comparisons(fit64, variables = variables64, newdata = dat64)
effects64

age64 <- avg_comparisons(fit64, variables = list(age = sd(dat64$age)),
                         newdata = dat64)
hypotheses(age64, hypothesis = difference ~ revpairwise)
```

**Example interpretations:** Women report about 0.99 more poor
mental-health days and 0.77 more poor physical-health days per month than men.
Although the effect of gender is about 0.22 larger for mental health, the
cross-outcome difference is not statistically significant.

Being married significantly reduces poor mental-health days by about 1.01,
whereas its effect on poor physical-health days is not statistically
significant. The direct comparison shows that the effect of marriage is
significantly larger for mental health than for physical health
(cross-model difference = -0.851, \(p < .01\)).

Similarly, the effect of age differs significantly across the outcomes:
aging is associated with fewer poor mental-health days but more poor
physical-health days.

### Ex 6.5 - Comparing marginal effects across different model types (ordinal vs nominal)

This example compares effects across two model types, an ordinal model and a
nominal model, with the same outcome and predictors. It also shows a custom
effect for a continuous predictor: a change in age from 20 to 30, with the
other covariates held at their means.

```{r example-6-5-models}
vars65 <- c("partyid5", "woman", "edyrs", "age", "parent", "married",
            "faminc", "employed", "region4", "year", "race")
dat65 <- subset(gss, year >= 2010)
dat65 <- dat65[complete.cases(dat65[vars65]), ]
stopifnot(nrow(dat65) == 8179)
nrow(dat65)
dat65 <- factorize(dat65, c("woman", "parent", "married", "race",
                            "employed", "region4", "year"))

party_levels <- sort(unique(dat65$partyid5))
party_labels <- c("Strong Democrat", "Democrat", "Independent",
                  "Republican", "Strong Republican")
dat65$party_ord <- ordered(dat65$partyid5, levels = party_levels, labels = party_labels)
dat65$party_nom <- factor(dat65$partyid5, levels = party_levels, labels = party_labels)

ordered65 <- MASS::polr(party_ord ~ age + I(age^2) + woman + edyrs + parent +
                          married + race + faminc + employed + region4 + year,
                        data = dat65, method = "logistic", Hess = TRUE)
nominal65 <- nnet::multinom(party_nom ~ age + I(age^2) + woman + edyrs + parent +
                             married + race + faminc + employed + region4 + year,
                           data = dat65, Hess = TRUE, trace = FALSE)
fit65 <- suest(ordered65, nominal65, model_names = c("Ordered", "Multinomial"))
```

It's important to be clear about what is being estimated here. The
`mecompare` Stata command compares age 20 with age 30 and holds the other
**model-matrix columns** at their means. That is not the same as averaging a
ten-year increase over everyone's observed age and covariates, and the
difference matters a lot here because age is included as a quadratic.

`newdata = "mean"` in `marginaleffects` doesn't reproduce it exactly either,
because it uses means for numeric variables but modes for categorical ones.
The helper included with `suest` reproduces Stata's `atmeans` calculation
exactly. [View the helper code](https://github.com/tdmize/suest/blob/main/inst/example-code/example-6-5-atmeans.R).

```{r example-6-5-effects}
source(system.file("example-code", "example-6-5-atmeans.R", package = "suest"))
results65 <- example65_atmeans(fit65, age_lo = 20, age_hi = 30)
results65$effects
results65$differences
```

**Example interpretation:** For someone who is 20 years old, the effect of a
ten-year increase in age differs significantly across the ordinal and nominal
models for three of the five outcome categories. For example, aging increases
the probability of identifying as a strong Democrat in both models, but the
increase is significantly larger in the nominal model. More strikingly, the
effects of age on identifying as a strong Republican run in opposite
directions across the two models.

### Ex 6.6 - Comparing marginal effects across different samples or groups

This example compares effects from the same binary logit model fitted to two
different samples, one from 1986 and one from 2016. The same approach works for
comparing groups, such as men and women. In R, the samples are selected with
`subset`, and `suest_newdata()` makes sure each marginal effect is averaged over
its own model's sample.

```{r example-6-6}
vars66 <- c("helpsickB", "polviews", "conserv", "faminc", "employed",
            "woman", "age", "college", "married", "parent", "race", "year")
dat66 <- gss[complete.cases(gss[vars66]), ]
dat66 <- factorize(dat66, c("conserv", "employed", "woman", "college",
                            "married", "parent", "race"))

model1986 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 1986)
model2016 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 2016)
stopifnot(nobs(model1986) == 1254, nobs(model2016) == 1670)
c(`1986` = nobs(model1986), `2016` = nobs(model2016))
fit66 <- suest(model1986, model2016, model_names = c("1986", "2016"))
fit66

effects66 <- avg_comparisons(fit66, variables = "conserv",
                             newdata = suest_newdata(fit66))
effects66
hypotheses(effects66, hypothesis = difference ~ revpairwise)
```

**Example interpretation:** Conservatives had significantly lower predicted
probabilities of saying that government should be responsible for providing
health care in both 1986 (marginal effect = -0.092) and 2016 (marginal effect =
-0.258). The gap between conservatives and nonconservatives increased over
time: the conservative effect is significantly more negative in 2016 than in
1986 (cross-model difference = 0.166, \(p < .001\)).

## Alternative fitting engines

`suest` handles the type of model and the package that fits it separately.
This lets you combine models fitted with different packages, as long as their
scores, parameter order, observations, and prediction scales are compatible.

### `glm2::glm2()`

```{r glm2-engine}
library(glm2)

standard <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
alternative <- glm2(am ~ wt + hp, family = binomial("logit"), data = dat)
engine_fit <- suest(standard, alternative, model_names = c("glm", "glm2"))
engine_fit
avg_comparisons(engine_fit, variables = "wt", newdata = dat)
```

### `ordinal::clm()`

`clm()` is supported for standard ordered logit and probit models: flexible
thresholds and no scale or nominal-effects formula.

```{r clm-engine}
library(ordinal)

housing <- MASS::housing[rep(seq_len(nrow(MASS::housing)), MASS::housing$Freq),
                         c("Sat", "Infl", "Type", "Cont")]
rownames(housing) <- NULL
housing$Sat <- ordered(housing$Sat, levels = c("Low", "Medium", "High"))

clm_model <- clm(Sat ~ Cont + Infl + Type, data = housing, link = "logit")
polr_model <- MASS::polr(Sat ~ Cont + Infl + Type, data = housing,
                         method = "logistic", Hess = TRUE)
ordinal_fit <- suest(clm_model, polr_model, model_names = c("clm", "polr"))
avg_comparisons(ordinal_fit, variables = "Cont", newdata = housing)
```

## For Stata users

The same framework is available in Stata through the
[`mecompare` command](https://www.trentonmize.com/software/mecompare). This R
package uses `marginaleffects` to calculate predictions and effects, while
`mecompare` does the same kinds of cross-model comparisons entirely in Stata.
