---
title: "Getting started with suest"
description: "Compare predictions and marginal effects across regression models in R."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with suest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

`suest()` combines two or more separately fitted models so you can test
whether predictions, marginal effects, or other quantities differ across the
models. It supplies the joint robust covariance matrix those tests need, and
leaves calculating and presenting the quantities themselves to
[`marginaleffects`](https://marginaleffects.com/).

## Functions at a glance

- `suest()` combines two or more fitted models and calculates their joint
  covariance matrix.
- `suest_newdata()` stacks the models' estimation samples so each model's
  effects can be averaged over its own sample.
- `suest_mi()` pools `suest()` results across multiply imputed datasets.
- The usual `marginaleffects` functions, including `predictions()`,
  `avg_comparisons()`, `avg_slopes()`, `hypotheses()`, and the plotting
  functions, work with the combined object.

## Installation and setup

```{r installation, eval=FALSE}
remotes::install_github("tdmize/suest")
```

```{r libraries}
library(suest)
library(marginaleffects)
```

## Basic workflow

```{r basic-workflow}
dat <- mtcars
dat$am <- factor(dat$am)

base <- glm(am ~ wt, family = binomial("logit"), data = dat)
adjusted <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
combined <- suest(base, adjusted, model_names = c("Base", "Adjusted"))
combined

effects <- avg_comparisons(combined, variables = "wt", newdata = dat)
effects
hypotheses(effects, hypothesis = difference ~ revpairwise)
```

The last command tests the difference between the two effects directly. That
is different from checking whether one effect is statistically significant and
the other is not.

## Supported models

### Single-level models

| Model | Fitted with |
|---|---|
| Linear regression | `lm()` |
| Binary logit, probit, and complementary log-log | `glm()`, `glm2::glm2()` |
| Ordered logit and probit | `MASS::polr()`, `ordinal::clm()` |
| Multinomial logit | `nnet::multinom()` |
| Poisson regression | `glm()`, `glm2::glm2()` |
| Negative binomial regression | `MASS::glm.nb()` |
| Zero-inflated Poisson and negative binomial | `pscl::zeroinfl()` |
| Other GLMs with identity, log, logit, probit, complementary log-log, or log-log links | `glm()`, `glm2::glm2()` |
| Tobit regression | `censReg::censReg()`; predictions are the latent mean |
| Interval, censored, and parametric survival regression | `survival::survreg()` with a common scale |
| Truncated regression | `truncreg::truncreg()` |
| Fractional response | `glm()` with `quasibinomial()`: logit, probit, complementary log-log, or log-log link |
| Beta regression | `betareg::betareg()` with a logit, probit, complementary log-log, or log-log link |
| Heteroskedastic probit and logit | `Rchoice::hetprob()` |
| Bivariate probit | `mvProbit::mvProbit()` with the same predictors in both equations |
| Instrumental-variable regression (two-stage least squares) | `fixest::feols()`, unweighted and without absorbed fixed effects |
| Instrumental-variable probit | `Rchoice::ivpml()`; predictions are average structural probabilities |
| Generalized ordered logit (gologit2), including partial proportional odds | `VGAM::vglm()` with `cumulative(reverse = TRUE)` |
| Heckman selection (maximum likelihood) | `sampleSelection::selection(method = "ml")` |

### Panel models

| Model | Fitted with |
|---|---|
| Linear fixed-, between-, and random-effects | `plm::plm()` with individual effects |
| Random-effects logit and probit | `pglm::pglm(model = "random", effect = "individual")` |
| Random-effects Poisson | `pglm::pglm(model = "random", effect = "individual", other = "sd")` |
| Fixed-effects Poisson | `fixest::fepois(y ~ x | id)` |
| Population-averaged (GEE) linear, logit, probit, complementary log-log, and Poisson | `geepack::geeglm()` with independence or exchangeable correlation |

### Multilevel models

| Model | Fitted with |
|---|---|
| Linear random-intercept | `nlme::lme(method = "ML")` |
| Random-intercept logit, probit, cloglog, Poisson, negative binomial, linear, and Gamma (log link) | `glmmTMB::glmmTMB()` |
| Random intercept and slope, `(1 + x | id)`, for the same families | `glmmTMB::glmmTMB()` |
| Linear random intercept or slope (ML) | `lme4::lmer(REML = FALSE)` |
| Random-intercept or slope logit, probit, cloglog, and Poisson | `lme4::glmer()`, any `nAGQ` |
| Random-intercept ordered logit and probit | `ordinal::clmm()` |

### Survey models

| Model | Fitted with |
|---|---|
| Linear regression | `survey::svyglm()` |
| Binary logit, probit, and cloglog | `survey::svyglm()` with `quasibinomial()` |
| Poisson | `survey::svyglm()` with `quasipoisson()` |

See [Survey data](#survey-data) below.

### Combining models

Models can use the same, partially overlapping, or completely different
samples. Any single-level models can be combined, including models for
different outcomes and models for categorical outcomes. Results on different
scales are labeled separately rather than treated as directly comparable.
Random-effects models (`nlme::lme()`, `pglm()`, `glmmTMB()`, lme4, `clmm()`) can be combined
with each other, including different families, and with linear, binary, count,
and ordered models, such as a logit model with a random-intercept logit model.
They must share one grouping variable, which the system is clustered on, and
must be unweighted. Other panel models (`plm()`, `geeglm()`) can be combined
only with models of the same kind. Offsets are supported, except in `pglm()`
and `glmmTMB()` models.

A few models have settings worth knowing about; see
[Notes on specific models](model-notes.html).

## Different samples or groups

When the models are fitted to different samples, such as separate groups,
use `suest_newdata()` so each model's effects are averaged over its own
sample:

```{r groups}
four <- lm(mpg ~ wt + hp, data = mtcars, subset = cyl == 4)
other <- lm(mpg ~ wt + hp, data = mtcars, subset = cyl != 4)
fit_groups <- suest(four, other, model_names = c("Four cylinders", "Other"))

effects_groups <- avg_comparisons(fit_groups, variables = "wt",
                                  newdata = suest_newdata(fit_groups))
effects_groups
hypotheses(effects_groups, hypothesis = difference ~ revpairwise)
```

## Cluster-robust standard errors

Supply a cluster variable to get cluster-robust standard errors for the
combined models:

```{r cluster-example, eval=FALSE}
fit_clustered <- suest(base, adjusted, cluster = "person_id")
avg_comparisons(fit_clustered, variables = "x", newdata = dat)
```

`cluster` takes one or more column names, or a list with one vector, matrix,
or data frame matched to each model's estimation sample. An observation that
appears in more than one model must have the same cluster ID in each.
Observations in completely different samples can belong to the same cluster,
in which case the covariance between models can be nonzero.

Panel models cluster on the panel identifier by default. A cluster you supply
must contain whole panels.

## Probability weights

`suest` supports probability weights for linear, binary logit and probit,
Poisson, negative binomial, ordered logit and probit, and multinomial logit
models. Fit each model with `weights=`, and tell `suest()` that they are
probability weights with `weight_type = "pweight"`:

```{r pweight-example}
set.seed(376)
weighted <- data.frame(x = rnorm(500), z = rnorm(500))
weighted$mediator <- 0.5 * weighted$x - 0.3 * weighted$z + rnorm(500)
weighted$pw <- exp(0.2 * weighted$z + 0.1 * rnorm(500))
weighted$y <- 1 + 0.7 * weighted$x - 0.4 * weighted$z +
  0.6 * weighted$mediator + rnorm(500)

base_w <- lm(y ~ x + z, weights = pw, data = weighted)
adjusted_w <- lm(y ~ x + z + mediator, weights = pw, data = weighted)
fit_w <- suest(base_w, adjusted_w, model_names = c("Base", "Adjusted"),
               weight_type = "pweight")
fit_w

effects_w <- avg_comparisons(fit_w, variables = "x", newdata = weighted, wts = "pw")
effects_w
hypotheses(effects_w, hypothesis = difference ~ revpairwise)
```

Weights must be positive and finite. An observation that appears in more than
one model must have the same weight in each; weights can differ for
observations in only one model, including completely different samples. When
the models have different estimation samples, `suest_newdata()` includes the
weights as `.suest_weight`:

```{r pweight-model-specific, eval=FALSE}
nd_w <- suest_newdata(fit_w)
avg_comparisons(fit_w, variables = "x", newdata = nd_w, wts = ".suest_weight")
```

## Survey data

`suest` supports linear, binary (logit, probit, cloglog), and Poisson models
fitted with `survey::svyglm()` from one common one-stage survey design, in any
combination:

```{r survey-example, eval=FALSE}
library(survey)
design <- svydesign(~psu, strata = ~strata, weights = ~w, data = dat)
m1 <- svyglm(yb ~ x, design, family = quasibinomial("logit"))
m2 <- svyglm(yb ~ x + z, design, family = quasibinomial("logit"))
fit <- suest(m1, m2, survey_design = design, observation_id = "id")
avg_slopes(fit, variables = "x", newdata = suest_newdata(fit), wts = ".suest_weight")
```

If either model uses a subset (domain) of the data, keep `design` as the full
sampled design. The joint covariance accounts for strata, PSUs, and optional
one-stage finite population corrections. Replicate weights, multistage
designs, calibration, and PPS designs are not supported.

## Multiple imputation

To combine results across imputed datasets, fit the same `suest` system in
each one. With `mice`, `with()` does this, and `marginaleffects` functions
then estimate within each imputation and pool the results with Rubin's
rules, including cross-model differences:

```{r mi-example, eval=FALSE}
analyses <- with(imp, suest(
  glm(y ~ x, family = binomial),
  glm(y ~ x + z, family = binomial),
  model_names = c("Base", "Adjusted")
))
avg_comparisons(analyses, variables = "x",
  hypothesis = difference ~ revpairwise)
```

`mice` warns "Large sample assumed." here. That is expected: SUEST reports
large-sample statistics, so the pooled degrees of freedom are Rubin's
large-sample values.

`suest_mi()` pools the coefficients and their joint covariance:

```{r mi-pool, eval=FALSE}
pooled <- suest_mi(analyses)
summary(pooled)
```

`marginaleffects` functions give the same results on `pooled` as on
`analyses`. A list of `suest()` results, one per imputation, or the result
of `with(mitools::imputationList(...), suest(...))` can also be passed to
`suest_mi()`; `marginaleffects` then warns that it could not recover the
original data from a `mids` object, which does not affect `suest()`
results.

## Examples

Six examples from Mize, Doan, and Long (2019), with code and output:

1. [Curvilinear effects and mediation/attenuation](mediation.html)
2. [Nested logit models and mediation/attenuation](nested-models.html)
3. [Alternative predictors](alternative-predictors.html)
4. [Different outcomes](different-outcomes.html)
5. [Different model types](model-types.html): ordinal versus nominal
6. [Separate samples or groups](groups.html)
