---
title: "Getting started with suest"
description: "Compare predictions and marginal effects across regression models in R."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with suest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
site <- identical(Sys.getenv("IN_PKGDOWN"), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = site,
                      message = FALSE, warning = FALSE)
options(width = 100)
```

`suest()` combines two separately fitted models and supplies the joint robust
covariance matrix needed to test whether predictions, marginal effects, or
other quantities differ across models. The package leaves calculation and
presentation of those quantities to
[`marginaleffects`](https://marginaleffects.com/).

The package implements the framework in Mize, Doan, and Long (2019). The six
replications below correspond to Examples 6.1–6.6 for the
[`mecompare` Stata command](https://www.trentonmize.com/software/mecompare).

For numeric predictors, a single numeric value in `variables` requests a
forward change of that amount. Thus, `list(income = sd(dat61$income))`
calculates the change from each observed income value to that value plus one
standard deviation. This differs from `"sd"`, which requests a centered
one-standard-deviation contrast around the regressor mean.

## Functions at a glance

- `suest()` combines two fitted models and calculates their joint covariance matrix.
- `suest_newdata()` stacks the two estimation samples so each model's effects
  can be averaged over its own sample.
- Standard `marginaleffects` functions—including `predictions()`,
  `avg_comparisons()`, `avg_slopes()`, `hypotheses()`, and the plotting
  functions—work with the combined object.

## Installation and setup

```{r installation, eval=FALSE}
remotes::install_github("tdmize/suest")
```

```{r libraries}
library(suest)
library(marginaleffects)
```

## Basic workflow

```{r basic-workflow}
dat <- mtcars
dat$am <- factor(dat$am)

base <- glm(am ~ wt, family = binomial("logit"), data = dat)
adjusted <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
combined <- suest(base, adjusted, model_names = c("Base", "Adjusted"))
combined

effects <- avg_comparisons(combined, variables = "wt", newdata = dat)
effects
hypotheses(effects, hypothesis = difference ~ revpairwise)
```

The final command tests the difference between the two effects directly. It
does not compare whether one effect is statistically significant while the
other is not.

## Supported models

| Model | Supported functions |
|---|---|
| Linear regression | `stats::lm()` |
| Binary logit and probit | `stats::glm()`, `glm2::glm2()` |
| Poisson regression | `stats::glm()`, `glm2::glm2()` |
| Negative binomial regression | `MASS::glm.nb()` |
| Ordered logit and probit | `MASS::polr()`, restricted `ordinal::clm()` |
| Multinomial logit | `nnet::multinom()` |

Models may use identical, partially overlapping, or disjoint samples. Supported
cross-family comparisons are logit–probit, logit–linear, probit–linear,
Poisson–negative binomial, and ordered logit–multinomial logit.

Nonunit weights and offsets are not yet supported. Bias-reduced,
adjusted-score, Firth, and penalized GLM fits are rejected because their
estimating equations differ from ordinary maximum-likelihood GLM scores.

## Replicating Mize, Doan, and Long (2019)

The examples use the public replication data for the
[`mecompare` Stata command](https://www.trentonmize.com/software/mecompare).
The data are downloaded once and reused below.

```{r replication-data}
library(haven)

ah <- zap_labels(read_dta("https://tdmize.github.io/data/data/ah4_cme.dta"))
gss <- zap_labels(read_dta("https://tdmize.github.io/data/data/gss_cme.dta"))

factorize <- function(data, variables) {
  data[variables] <- lapply(data[variables], factor)
  data
}
```

### Ex 6.1 - Marginal effects to summarize curvilinear relationships and test mediation

This example uses a linear regression model with a nonlinear effect represented
by a polynomial: both income and income squared are included in the model. We
test whether adding the mediator, job satisfaction, reduces the average effect
of income, which would suggest mediation. The example also shows how to request
a custom amount of change for a continuous predictor—in this case, a
one-standard-deviation increase in income.

```{r example-6-1}
vars61 <- c("depsympB", "income", "inc10", "age", "woman", "race", "college", "jobsat")
dat61 <- ah[complete.cases(ah[vars61]), ]
stopifnot(nrow(dat61) == 4307)
nrow(dat61)
dat61 <- factorize(dat61, c("woman", "race", "jobsat"))

base61 <- lm(depsympB ~ income + I(income^2) + age + woman + race, data = dat61)
mediator61 <- lm(depsympB ~ income + I(income^2) + age + woman + race + jobsat,
                 data = dat61)
fit61 <- suest(base61, mediator61, model_names = c("Base", "Mediator"))

effects61 <- avg_comparisons(fit61,
  variables = list(income = sd(dat61$income)), newdata = dat61)
effects61
hypotheses(effects61, hypothesis = difference ~ revpairwise)
```

**Example interpretation for the effect of income:** In model 1, the
marginal effect is -0.816, indicating that a standard-deviation increase in
income is associated with about 0.8 fewer depressive symptoms. The effect is
reduced to -0.648 after accounting for job satisfaction. The direct
cross-model test shows that accounting for job satisfaction decreases the
average effect of income by 0.168, a statistically significant reduction
(\(p < .001\)) that is consistent with mediation.

### Ex 6.2 - Comparing marginal effects across nested logit models

This example uses binary logit models and examines how the effect of college
changes as additional variables are added to the model.

```{r example-6-2}
vars62 <- c("vhappy", "college", "wages", "occprest", "age", "married",
            "parent", "woman", "conserv", "reltrad", "year", "employed")
dat62 <- subset(gss, year >= 2000 & employed == 1)
dat62 <- dat62[complete.cases(dat62[vars62]), ]
stopifnot(nrow(dat62) == 9216)
nrow(dat62)
dat62 <- factorize(dat62, c("college", "married", "parent", "woman", "conserv",
                            "reltrad", "year"))

model62a <- glm(vhappy ~ college, family = binomial("logit"), data = dat62)
model62b <- glm(vhappy ~ college + married + parent + woman + conserv + reltrad +
                  year + age + I(age^2), family = binomial("logit"), data = dat62)
fit62 <- suest(model62a, model62b, model_names = c("Model 1", "Model 2"))

effects62 <- avg_comparisons(fit62, variables = "college", newdata = dat62)
effects62
hypotheses(effects62, hypothesis = difference ~ revpairwise)
```

**Example interpretation for the effect of college:** On average, the
probability of being very happy is 0.072 higher for people with a college
degree than for those without one (\(p < .001\)). Model 2 adds demographic
controls, reducing the average marginal effect of college to 0.060, which
remains statistically significant. A direct test shows that adding the
controls decreases the effect of college by 0.012 (\(p < .01\)).

### Ex 6.3 - Comparing marginal effects using alternative predictors

This example uses two binary logit models with alternative predictors. The
models use two different ways to measure sexuality: sexual behavior
(`sexbehav`) in one model and sexual identity (`sexident`) in the other.

```{r example-6-3-models}
vars63 <- c("samesexB", "sexident", "sexbehav", "college",
            "woman", "race", "age", "year")
dat63 <- gss[complete.cases(gss[vars63]), ]
stopifnot(nrow(dat63) == 4921)
nrow(dat63)
dat63 <- factorize(dat63, c("sexident", "sexbehav", "college", "woman", "race", "year"))

behavior63 <- glm(samesexB ~ sexbehav + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
identity63 <- glm(samesexB ~ sexident + woman + college + age + race + year,
                  family = binomial("logit"), data = dat63)
fit63 <- suest(behavior63, identity63, model_names = c("Behavior", "Identity"))

effects63 <- avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"), newdata = dat63)
effects63
```

Because each predictor appears in only one component model,
`marginaleffects` reports zero and `NA` for structurally absent effects. The
nonzero rows give the effects for each operationalization.

```{r example-6-3-contrasts}
alternative_predictor_hypothesis <- function(x) {
  group <- as.character(x$group)
  term <- as.character(x$term)
  behavior <- x$estimate[term == "sexbehav" & group == "Behavior"]
  identity <- x$estimate[term == "sexident" & group == "Identity"]

  data.frame(
    term = c("Behavior bisexual - Identity bisexual",
             "Behavior gay - Identity gay",
             "(Behavior gay - bisexual) - (Identity gay - bisexual)"),
    estimate = c(behavior[1] - identity[1], behavior[2] - identity[2],
                 (behavior[2] - behavior[1]) - (identity[2] - identity[1])))
}

avg_comparisons(fit63,
  variables = list(sexbehav = "reference", sexident = "reference"),
  newdata = dat63, hypothesis = alternative_predictor_hypothesis)
```

**Example interpretations:** The contrasts between heterosexual and
gay/lesbian respondents and between bisexual and gay/lesbian respondents do
not differ significantly across the two models. However, the difference
between heterosexual and bisexual respondents is significantly larger when
sexuality is measured using identity than when it is measured using behavior.

### Ex 6.4 - Comparing marginal effects across different outcomes

This example compares effects across different dependent variables. Two count
outcomes—poor mental-health days and poor physical-health days—are estimated
using negative-binomial models, with effects calculated on the predicted-rate
scale. Effects and cross-model comparisons are calculated for all predictors.
To reproduce the `mecompare` Stata example exactly, the common sample also
excludes observations missing `reltrad`, even though `reltrad` is not included
in either displayed model.

```{r example-6-4}
vars64 <- c("mntlhlth", "physhlth", "woman", "married", "age",
            "faminc", "race", "college", "parent", "reltrad")
dat64 <- gss[complete.cases(gss[vars64]), ]
stopifnot(nrow(dat64) == 5062)
nrow(dat64)

dat64 <- factorize(dat64, c("woman", "married", "parent", "college", "race", "year"))

mental64 <- MASS::glm.nb(mntlhlth ~ woman + married + parent + college + age +
                           faminc + race + year, data = dat64)
physical64 <- MASS::glm.nb(physhlth ~ woman + married + parent + college + age +
                             faminc + race + year, data = dat64)
fit64 <- suest(mental64, physical64, model_names = c("Mental", "Physical"))

variables64 <- list(woman = "reference", married = "reference", parent = "reference",
  college = "reference", age = sd(dat64$age), faminc = sd(dat64$faminc),
  race = "reference", year = "reference")
effects64 <- avg_comparisons(fit64, variables = variables64, newdata = dat64)
effects64

age64 <- avg_comparisons(fit64, variables = list(age = sd(dat64$age)),
                         newdata = dat64)
hypotheses(age64, hypothesis = difference ~ revpairwise)
```

**Example interpretations:** Women report about 0.99 more poor
mental-health days and 0.77 more poor physical-health days per month than men.
Although the effect of gender is about 0.22 larger for mental health, the
cross-outcome difference is not statistically significant.

Being married significantly reduces poor mental-health days by about 1.01,
whereas its effect on poor physical-health days is not statistically
significant. The direct comparison shows that the effect of marriage is
significantly larger for mental health than for physical health
(cross-model difference = -0.851, \(p < .01\)).

Similarly, the effect of age differs significantly across the outcomes:
aging is associated with fewer poor mental-health days but more poor
physical-health days.

### Ex 6.5 - Comparing marginal effects across different model types (ordinal vs nominal)

This example compares effects across two model types: an ordinal model and a
nominal model. The models are otherwise identical, using the same outcome and
predictors. It also illustrates a custom effect for a continuous predictor:
a change in age from 20 to 30, with the other covariates held at their means.

```{r example-6-5-models}
vars65 <- c("partyid5", "woman", "edyrs", "age", "parent", "married",
            "faminc", "employed", "region4", "year", "race")
dat65 <- subset(gss, year >= 2010)
dat65 <- dat65[complete.cases(dat65[vars65]), ]
stopifnot(nrow(dat65) == 8179)
nrow(dat65)
dat65 <- factorize(dat65, c("woman", "parent", "married", "race",
                            "employed", "region4", "year"))

party_levels <- sort(unique(dat65$partyid5))
party_labels <- c("Strong Democrat", "Democrat", "Independent",
                  "Republican", "Strong Republican")
dat65$party_ord <- ordered(dat65$partyid5, levels = party_levels, labels = party_labels)
dat65$party_nom <- factor(dat65$partyid5, levels = party_levels, labels = party_labels)

ordered65 <- MASS::polr(party_ord ~ age + I(age^2) + woman + edyrs + parent +
                          married + race + faminc + employed + region4 + year,
                        data = dat65, method = "logistic", Hess = TRUE)
nominal65 <- nnet::multinom(party_nom ~ age + I(age^2) + woman + edyrs + parent +
                             married + race + faminc + employed + region4 + year,
                           data = dat65, Hess = TRUE, trace = FALSE)
fit65 <- suest(ordered65, nominal65, model_names = c("Ordered", "Multinomial"))
```

The estimand here is important. The `mecompare` Stata command compares
age 20 with age 30 and holds the other **model-matrix columns** at their means.
This is not the same as averaging a ten-year forward change over observed ages
and covariates. The distinction is especially consequential because age enters
quadratically.

It is also not reproduced exactly by `newdata = "mean"` in `marginaleffects`,
which uses means for numeric variables but modes for categorical variables.
The helper supplied with `suest` reproduces Stata's model-matrix `atmeans`
calculation exactly. [View the helper code](https://github.com/tdmize/suest/blob/main/inst/example-code/example-6-5-atmeans.R).

```{r example-6-5-effects}
source(system.file("example-code", "example-6-5-atmeans.R", package = "suest"))
results65 <- example65_atmeans(fit65, age_lo = 20, age_hi = 30)
results65$effects
results65$differences
```

**Example interpretation:** For someone who is 20 years old, the effect of a
ten-year increase in age differs significantly across the ordinal and nominal
models for three of the five outcome categories. For example, aging increases
the probability of identifying as a strong Democrat in both models, but the
increase is significantly larger in the nominal model. More strikingly, the
effects of age on identifying as a strong Republican run in opposite
directions across the two models.

### Ex 6.6 - Comparing marginal effects across different samples or groups

This example compares effects from separate models estimated on different
samples: one sample from 1986 and another from 2016. The models are otherwise
identical, using binary logit with the same outcome and predictors. The same
approach can be used to compare models for different groups, such as men and
women. In R, the samples are selected with `subset`, and `suest_newdata()`
ensures that each marginal effect is averaged over the appropriate estimation
sample.

```{r example-6-6}
vars66 <- c("helpsickB", "polviews", "conserv", "faminc", "employed",
            "woman", "age", "college", "married", "parent", "race", "year")
dat66 <- gss[complete.cases(gss[vars66]), ]
dat66 <- factorize(dat66, c("conserv", "employed", "woman", "college",
                            "married", "parent", "race"))

model1986 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 1986)
model2016 <- glm(helpsickB ~ conserv + faminc + employed + woman + age +
                   college + married + parent + race, family = binomial("logit"),
                 data = dat66, subset = year == 2016)
stopifnot(nobs(model1986) == 1254, nobs(model2016) == 1670)
c(`1986` = nobs(model1986), `2016` = nobs(model2016))
fit66 <- suest(model1986, model2016, model_names = c("1986", "2016"))
fit66

effects66 <- avg_comparisons(fit66, variables = "conserv",
                             newdata = suest_newdata(fit66))
effects66
hypotheses(effects66, hypothesis = difference ~ revpairwise)
```

**Example interpretation:** Conservatives had significantly lower predicted
probabilities of saying that government should be responsible for providing
health care in both 1986 (marginal effect = -0.092) and 2016 (marginal effect =
-0.258). The gap between conservatives and nonconservatives increased over
time: the conservative effect is significantly more negative in 2016 than in
1986 (cross-model difference = 0.166, \(p < .001\)).

## Alternative fitting engines

The statistical model family and fitting engine are handled separately. This
allows comparisons across packages when likelihood scores, parameter ordering,
observation alignment, and prediction scales are compatible.

### `glm2::glm2()`

```{r glm2-engine}
library(glm2)

standard <- glm(am ~ wt + hp, family = binomial("logit"), data = dat)
alternative <- glm2(am ~ wt + hp, family = binomial("logit"), data = dat)
engine_fit <- suest(standard, alternative, model_names = c("glm", "glm2"))
engine_fit
avg_comparisons(engine_fit, variables = "wt", newdata = dat)
```

### `ordinal::clm()`

`clm()` is supported for proportional-effects ordered logit and probit models
with flexible thresholds and no scale or nominal-effects formula.

```{r clm-engine}
library(ordinal)

housing <- MASS::housing[rep(seq_len(nrow(MASS::housing)), MASS::housing$Freq),
                         c("Sat", "Infl", "Type", "Cont")]
rownames(housing) <- NULL
housing$Sat <- ordered(housing$Sat, levels = c("Low", "Medium", "High"))

clm_model <- clm(Sat ~ Cont + Infl + Type, data = housing, link = "logit")
polr_model <- MASS::polr(Sat ~ Cont + Infl + Type, data = housing,
                         method = "logistic", Hess = TRUE)
ordinal_fit <- suest(clm_model, polr_model, model_names = c("clm", "polr"))
avg_comparisons(ordinal_fit, variables = "Cont", newdata = housing)
```

## For Stata users

The same framework is available in Stata through the
[`mecompare` Stata command](https://www.trentonmize.com/software/mecompare).
The R package uses `marginaleffects` for predictions and effects, while the
`mecompare` Stata command provides a Stata-native workflow for the same class
of cross-model comparisons.
